Covariate analysis of UFL-Mayo-ISB RNASeq temporal cortex

### 2016-01-11 code/Rmd/Mayo_TemporalCortex_RNASeq.Rmd
Added markdown to compute covariates for Mayo temporal cortex

---
title: "Covariate analysis of UFL-Mayo-ISB RNASeq"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./Mayo_Cerebellum_RNASeq.Rmd",
                                 parentId ="syn5570325",
                                 entityName = 'UFL_Mayo_ISB Cerebellum RNASeq')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570325';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of UFL_Mayo_ISB cerebellum RNASeq data';

thisFileName <- 'Mayo_Cerebellum_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn5201012';
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- synGet(COUNT_ID)
COUNT <- read.table(COUNT_OBJ@filePath,header=T,sep='\t',check.names = F)

# Get clinical metadata
METADATA_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA_OBJ <- synGet(METADATA_ID)
METADATA <- read.table(METADATA_OBJ@filePath,sep=',',header=T, stringsAsFactors=F)
```

@@

```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))

@@

# Match covariates to expression data
rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[rownames(METADATA) %in% indToRetain,]

@@

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
FactorCovariates <- c("Source", "Diagnosis",  "Sex",  "ApoE", "Flowcell")
ContCovariates <- c("RIN", "AgeAtDeath")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$SampleID

@@

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, METADATA, by="SampleID")

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum RNASeq', parentId = parentId)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

>>>>

---
title: "Covariate analysis of UFL-Mayo-ISB RNASeq temporal cortex"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./Mayo_TemporalCortex_RNASeq.Rmd",
                                 parentId ="syn5570325",
                                 entityName = 'UFL_Mayo_ISB Temporal Cortex RNASeq')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570325';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of UFL_Mayo_ISB temporal cortex RNASeq data';

thisFileName <- 'Mayo_TemporalCortex_RNASeq.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn4650257';
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- synGet(COUNT_ID)
COUNT <- read.table(COUNT_OBJ@filePath,header=T,sep=' ',check.names = F)

# Get clinical metadata
METADATA_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA_OBJ <- synGet(METADATA_ID)
METADATA <- read.table(METADATA_OBJ@filePath,sep=',',header=T, stringsAsFactors=F)
```

@@

```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$ID, colnames(COUNT))

@@

# Match covariates to expression data
rownames(METADATA) = METADATA$ID
METADATA = METADATA[rownames(METADATA) %in% indToRetain,]

@@

### Covariate clustering
Determine relationship between covariates. 
```{r covariates.clustering, cache=TRUE}
FactorCovariates <- c("Source", "Diagnosis",  "FLOWCELL")
ContCovariates <- c("RIN", "AgeAtDeath")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$ID

@@

# Plot first 2 PCs
plotdata <- data.frame(ID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, METADATA, by="ID")

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Temporal Cortex RNASeq', parentId = parentId)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'ID')
write.table(COVARIATES, file = 'covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
